---
title: "Group 11 Project"
date: ' `r Sys.Date()`'
output:
  html_document:
    df_print: paged
---
Campaign Proposal <br>
Class: Data Mining (4080-001) <br>
Authors: An Nguyen, Avneet Dharni, Katie Dang, Roan Zappanti, Tatum Arey 
<br><br>
  
## Business Objective

Evaluate success factors and areas for growth for Campaign A, B, and C

* Who are we serving?
* What are we serving?
* Where should we serve it?

## Load the Packages & Get Data
```{r Loading Packages & Data}
library(tidyverse)
library(lubridate)
library(completejourney)
library(RColorBrewer)
library(dplyr)


c(promotions, transactions) %<-% get_data(which = 'both', verbose = FALSE)
```

## Initial Observations

We observed that Campaign A is leading the pack in sales. Both Campaign's B & C have room for improvement however especially C. For the majority of our analysis, we will hone in on understanding the attributes of a successful campaign to boost sales of future ones.


```{r Total Sales}
df2 <- df %>%
  group_by(campaign_type) %>%
  summarise(total_sales_campaign_type = sum(sales_value, na.rm = TRUE))

df2 %>% 
  ggplot(aes(x= campaign_type, y= total_sales_campaign_type, fill= campaign_type))+
  geom_bar(stat = "identity", show.legend = FALSE)+
  scale_fill_brewer(palette = "Oranges")+
  scale_y_continuous( name = "Sales Value of each Campaign Type", labels = scales::dollar)+
  scale_x_discrete(name= "Campaign Type")+
  labs( title = "Total Sales Value of Each Campaign Type",
       caption = "http://https://github.com/bradleyboehmke/completejourney")
```




#### Observation 1: Who are we serving?

Income is a large deciding factor for each campaign's success. The *50-74K income range* has the most counts of sales for each campaign type. We recommend that Regork continues to serve this demographic, while also innovating solutions to reach those missed markets.

```{r echo=TRUE, warning=FALSE, message=FALSE}

demographics %>%
  inner_join(transactions) %>%
  inner_join(campaigns) %>%
  inner_join(campaign_descriptions) %>%
  filter(campaign_type %in% c("Type B", "Type C")) %>%
  group_by(campaign_type) %>%
  count(income, campaign_type) %>%
  mutate(pct = n/sum(n)) %>%
  ggplot(aes(reorder(income,n), y = n, fill = campaign_type)) +
    geom_bar(stat = "identity") +
    theme(axis.title.y = element_blank()) +
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        plot.caption = element_text(size = 8, margin = margin(t = 10), color = "grey70", hjust = 0)) +
    scale_fill_brewer(palette = "Oranges") +
    geom_text(aes(label = round(pct, 2)), hjust = -0.1, position = position_stack(vjust = 0.01),
                     color = "black", check_overlap = TRUE) +
    scale_y_continuous(labels = scales::number) +
    coord_flip() +
  labs(
    title = "Income Distribution Among Campaign B & C",
    x = "Income range",
    y = "Count",
    caption = "Source: R data package that provides access to data in the Complete Journey package \n provided by 84.51",
    fill = "Campaign type")
```


<br>

## Observation 2: What are we serving?

As we originally stated, Campaign's A & B were shown to yield the highest sales value. Could what we are serving effect that? It appears that A & B offer almost *double* the amount of products as C. We recommend that we increase the number of products offered in each campaign for a wider reach to consumers.

```{r}
df <-transactions %>%
  left_join(products) %>%
  left_join(demographics) %>%
  full_join(campaigns) %>%
  full_join(campaign_descriptions)

  
# This graph show how many products is applied for each type of campaign
#This is the reason why campaign A has the most sales value 
#(df2 is the sales value of each campaign type)
df1 <- df %>%
  group_by(campaign_type) %>%
  count(product_id) %>%
  summarise(total_products= sum(n))

df1 %>% ggplot(aes(x= campaign_type,y=total_products)) +
  geom_segment( aes(x= campaign_type,xend= campaign_type, y=0, yend= total_products, color= "Orange"))+
  geom_point(color= 'Orange', size = 3) + 
  scale_y_continuous(trans= 'log2',name = "Number of products") +
  scale_x_discrete(name= "Campaign Type")+
  labs( title = "Number of products in each campaign type",
        subtitle = "This graph shows the number of product is applied in each campaign type",
       caption = "http://https://github.com/bradleyboehmke/completejourney")


df2<- df %>%
  group_by(campaign_type) %>%
summarise(total_sales_department= sum(sales_value))
```



<br>

## Observation 3: Where are we serving it?

0 & 7 are the top display location for Campaign's A & B. We recommend that Regork rotate the products through display locations 0 and 7.

```{r Defining Data & Variables, echo = FALSE, message= FALSE}
plot_data <- promotions %>% 
    inner_join(products, by = "product_id") %>% 
    inner_join(coupons, by = "product_id") %>% 
    inner_join(campaign_descriptions, by = "campaign_id") %>%
    select(campaign_type, display_location, product_id) %>% 
    group_by(campaign_type, display_location) %>% 
    summarize(distinct = n_distinct(product_id)) %>% 
    arrange(campaign_type, desc(distinct))

location_plot1 <- NA
location_plot2 <- NA
```

```{r Create the Function, echo = false}
# This function intakes the campaign type (ex: 'Type A') and outputs two ranked graphs regarding location counts by distinct product id

campaign_locations <- function(campaign_tp, messages = TRUE){
  
    
  location_plot1 <<- plot_data %>%
    filter(campaign_type == campaign_tp) %>% 
    ggplot(aes(x = reorder(display_location, -distinct), 
                y = distinct,
               fill = reorder(display_location, -distinct))) + 
      geom_bar(stat = "identity", show.legend = FALSE) +
      scale_y_continuous() +
      scale_fill_grey() +
      xlab("Display Locations") +
      ylab("Distinct Product Count") +
      labs( title = paste("# of unique products per display location"),
            subtitle = paste("For", campaign_tp))
    
  location_plot2 <<- plot_data %>% 
    filter(campaign_type == campaign_tp) %>% 
    ggplot(aes(x = distinct, 
               y = reorder(display_location, distinct),
               col = -distinct)) + 
      geom_point(show.legend = FALSE, size = 5) +
      scale_fill_distiller(type = "seq", palette = "Blues") +      
      xlab("Distinct Product Count") +
      ylab("Display Location") +
      labs( title = paste("# of unique products per display location"),
            subtitle = paste("For", campaign_tp))
  
  message("Function complete")
}  

```

```{r Run the Function, messages = FALSE}
# Running the 'campgain_locations()' function and displaying outputs
cmpgns <- c("Type A","Type B","Type C")
for(i in cmpgns){
  campaign_locations(i)
  print(location_plot1)
  print(location_plot2)
  rm(location_plot1)
  rm(location_plot2)
}

rm(cmpgns)
rm(i)
rm(plot_data)
```
