---
title: "Module 12 Lab"
author: "Roan Zappanti"
output: html_notebook
---

<br><br>
```{r}
### Lab Sections {.tabset .tabset-pills}
```

#### Part One: Tuning Regularized Regression

Managing global settings:
```{r Global-Setup, setup, include = TRUE}
knitr::opts_chunk$set(
  fig.width = 6, fig.height = 6,
  warning = FALSE, error = FALSE, message = FALSE,
  include = TRUE, echo = TRUE, strip.white = TRUE, highlight = TRUE,
  results = 'hold'
)
```


Loading Data:
```{r Libraries-and-Data}
library(tidymodels)
library(tidyverse)
library(vip)

path <- here::here("Data","boston.csv")
boston <- read_csv(path)
```

---

<br>
1. Split the data into a training set and test set using a 70-30% split. Be sure to include the
set.seed(123) so that your train and test sets are the same size as mine.

2. Create a recipe that will model cmedv as a function of all
predictor variables. Apply the following feature engineering steps
in this order:

    a. Normalize all numeric predictor variables using a Yeo-Johnson
    transformation
    b. Standardize all numeric predictor variables

3. Create a 5-fold cross validation resampling object.

4. Create a regularized regression model object that:

    a. Contains tuning placeholders for the mixture and penalty
      arguments
    b. Sets the engine to use the glmnet package.

5. Create our hyperparameter search grid that:
    
    a. Searches across default values for mixture
    b. Searches across values ranging from -10 to 5 for penalty
    c. Will search across 10 values for each of these
    hyperparameters   (levels)\n

6. Creates a workflow object that combines our recipe object with 
our model object.

7. Performs a hyperparameter search.

8. Assesses the results.


```{r , message = FALSE}
# Step 1. split our data
set.seed(123)
split <- initial_split(boston, prop = 0.7, strata = cmedv)
boston_train <- training(split)
boston_test <- testing(split)

boston_recipe <- recipe(cmedv ~ ., data = boston_train) %>%
step_YeoJohnson(all_numeric_predictors()) %>%
step_normalize(all_numeric_predictors())

# Step 3. create resampling object
set.seed(123)
kfolds <- vfold_cv(boston, v = 5, strata = cmedv)

# Step 4. create our model object
reg_mod <- linear_reg(mixture = tune(), penalty = tune()) %>%
set_engine("glmnet")

# Step 5. create our hyperparameter search grid
reg_grid <- grid_regular(mixture(), penalty(range = c(-10, 5)), levels = 10)

# Step 6. create our workflow object
boston_wf <- workflow() %>%
add_recipe(boston_recipe) %>%
add_model(reg_mod)

# Step 7. perform hyperparamter search
tuning_results <- boston_wf %>%
tune_grid(resamples = kfolds, grid = reg_grid)

# Step 8. assess results
tuning_results %>%
collect_metrics() %>%
filter(.metric == "rmse") %>%
arrange(mean)

autoplot(tuning_results)
```
Assess this plot regarding our hyperparameter search results. What do the results tell you? Does the amount of regularization (size of the penalty) have a larger influence on the RMSE or does the type of penalty applied (mixture) have a larger influence?
<br><br>
**Answer**: While the mixture has a slight effect on RMSE, penalty size maintains a far larger effect on RMSE. 

---
Now, fill in the blanks below to:

1. finalize our workflow object with the optimal hyperparameter values
2. fit our final workflow object across the full training set data, and
3. plot the top 10 most influential features.

```{r}
# Step 1. finalize our workflow object with the optimal hyperparameter values
best_hyperparameters <- select_best(tuning_results, metric = "rmse")

final_wf <- workflow() %>%
add_recipe(boston_recipe) %>%
add_model(reg_mod) %>%
finalize_workflow(best_hyperparameters)

# Step 2. fit our final workflow object across the full training set data
final_fit <- final_wf %>%
fit(data = boston_train)

# Step 3. plot the top 10 most influential features
final_fit %>%
extract_fit_parsnip() %>%
vip()
```
We see that lstat is the most influential predictor variable followed by dis and rm.

#### Part 2: Tuning a Regularized Classification Model

Your objective is to tune a regularized logistic regression model to find the hyperparameter values that
maximize the AUC model metric. Using the code chunk below, fill in the blanks to:

1. Split the data into a training set and test set using a 70-30% split. Be sure to include the set.seed(123) so that your train and test sets are the same size as mine.

2. Create a recipe that will model type as a function of all predictor variables. Apply the following feature engineering steps in this order:
   
    a. Normalize all numeric predictor variables using a
    Yeo-Johnson transformation
    b. Standardize all numeric predictor variables

3. Create a 5-fold cross validation resampling object.

4. Create a regularized logistic regression model object that:
    
    a. Contains tuning placeholders for the mixture and penalty
    arguments
    b. Sets the engine to use the glmnet package.
    c. Sets the mode to be a classification model.

5. Create our hyperparameter search grid that:
    
    a. Searches across default values for mixture
    b. Searches across values ranging from -10 to 5 for penalty
    c. Will search across 10 values for each of these
    hyperparameters (levels)

6. Creates a workflow object that combines our recipe object with our model object.

7. Performs a hyperparameter search.

8. Assesses the results.
```{r Gathering-Data}
# install.packages("kernlab") if you don't have this package installed
library(kernlab)
data(spam)

# Splitting the data
set.seed(123)
split <- initial_split(data = spam, prop = 0.7, strata = type)
spam_train <- training(split)
spam_test <- testing(split)

# Spam recipe
spam_recipe <- recipe(type ~ ., data = spam_train) %>% 
  step_YeoJohnson(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors())

# Creating CV object
set.seed(123)
kfolds <- vfold_cv(data = spam_train, v = 5, strata = type)

# Creating regularized logistic regression model with tuning params
logit_mod <- logistic_reg(mode = "classification", 
                          penalty = tune(),
                          mixture = tune(),
                          engine = "glmnet")

# Hyperparameter grid search for penalty and mixture
logit_grid <- grid_regular( mixture(), 
                           penalty(range = c(-10, 5)),
                           levels = 10)

# Create workflow and tune the parameters
spam_wf <- workflow() %>% 
  add_recipe(spam_recipe) %>% 
  add_model(logit_mod)

tuning_results <- spam_wf %>% 
  tune_grid(resamples = kfolds, grid = logit_grid)

tuning_results %>% 
  collect_metrics() %>% 
  filter(.metric == "roc_auc") %>%
  arrange(desc(mean))

#Plotting the results
autoplot(tuning_results)

```
Based on the graphed results of the grid search for the "penalty" and "mixture" parameters, the penalty size plays a greater role on the Area Under the Curve. All proportions of ridge vs lasso lead to a lower ROC after hitting a penalty size of ~ 0.0.

Now, fill in the blanks below to:

  1. finalize our workflow object with the optimal hyperparameter values,
  2. fit our final workflow object across the full training set data, and
  3. plot the top 10 most influential features.
```{r}

# Step 1. finalize our workflow object with the optimal hyperparameter values
best_hyperparameters <- select_best(tuning_results, metric = "roc_auc")

final_wf <- workflow() %>%
add_recipe(spam_recipe) %>%
add_model(logit_mod) %>%
finalize_workflow(best_hyperparameters)

# Step 2. fit our final workflow object across the full training set data
final_fit <- final_wf %>%
fit(data = spam_train)

# Step 3. plot the top 10 most influential features
final_fit %>%
extract_fit_parsnip() %>%
vip()
```
#### Part 3: Tuning a MARS Model

For this part of the lab we’ll continue using the kernlab::spam data set. Your objective is to tune a MARS classification model to find the hyperparameter values that maximize the AUC model metric. Using the code chunk below, fill in the blanks to:

1. Split the data into a training set and test set using a 70-30% split. Be sure to include the
  set.seed(123) so that your train and test sets are the same size as mine.
  
2. Create a recipe that will model type as a function of all predictor variables. Apply the following
  feature engineering steps in this order:
    
    a. Normalize all numeric predictor variables using a Yeo-Johnson transformation
    b. Standardize all numeric predictor variables

3. Create a 5-fold cross validation resampling object.

4. Create a MARS model object that:
    
    a. Contains tuning placeholders for the num_terms and prod_degree arguments
    b. Sets the mode to be a classification model.

5. Create our hyperparameter search grid that:
    
    a. Searches across values ranging from 1 to 30 for num_terms
    b. Searches across default values for prod_degree
    c. Will search across 25 values for each of these hyperparameters (levels)

6. Creates a workflow object that combines our recipe object with our model object.

7. Performs a hyperparameter search.

8. Assesses the results.

```{r}

# Step 4: create ridge model object
mars_mod <- mars(num_terms = tune(), prod_degree = tune()) %>%
set_mode("classification")

# Step 5. create our hyperparameter search grid
mars_grid <- grid_regular(num_terms(range = c(1,30)), prod_degree(), levels = 25)

# Step 6: create workflow object to combine the recipe & model
spam_wf <- workflow() %>%
add_recipe(spam_recipe) %>%
add_model(mars_mod)

# Step 7. perform hyperparamter search
tuning_results <- spam_wf %>%
tune_grid(resamples = kfolds, grid = mars_grid)

# Step 8. assess results
tuning_results %>%
collect_metrics() %>%
filter(.metric == "roc_auc") %>%
arrange(desc(mean))

autoplot(tuning_results)
```
Based on the graphs above, number of model terms (with a platuau around n=10) has a higher affect on roc_auc than the degree of interaction. 

```{r}
# Step 1. finalize our workflow object with the optimal hyperparameter values
best_hyperparameters <- select_best(tuning_results, metric = "roc_auc")

final_wf <- workflow() %>%
add_recipe(spam_recipe) %>%
add_model(mars_mod) %>%
finalize_workflow(best_hyperparameters)

# Step 2. fit our final workflow object across the full training set data
final_fit <- final_wf %>%
fit(data = spam_train)

# Step 3. plot the top 10 most influential features
final_fit %>%
extract_fit_parsnip() %>%
vip()
```
